---
title: "Quiz1"
author: "Jens Schreiber"
date: "5. Oktober 2015"
output: html_document
---

Question 1:

Consider the following code for the cars data set
```{r, cache=TRUE}
  library(manipulate)
  myPlot <- function(s) {
    plot(cars$dist - mean(cars$dist), cars$speed - mean(cars$speed))
    abline(0, s)
  }
```
This function plots distance versus speed, each de-meaned and an associated line of slope s. Which of the following code will make a manipulate plot that creates a slider for the slope?
```{r, cache=TRUE}
  #manipulate(myPlot(s), s = slider(0, 2, step = 0.1))
```

Question 2
Which of the following code uses the rCharts package to create a sortable and searchable data table for the airquality data set? Assume the rCharts package and the airquality data set have already been loaded into R.

```{r, cache=TRUE}
  library(rCharts)
  require(graphics)
  dTable(airquality, sPaginationType = "full_numbers")
```

Question 3
A basic shiny data product requires:
--> A ui.R and server.R file or a A server.R file and a directory called www containing the relevant html files.

Question 4:
What is incorrect about the followig syntax in ui.R?
```{r, cache=TRUE}
  #library(shiny)
  #shinyUI(pageWithSidebar(
  #  headerPanel("Data science FTW!"),
  #  sidebarPanel(
  #  h2('Big text')
  #  h3('Sidebar')
  #),
  #mainPanel(
  #    h3('Main Panel text')
  #)
#  ))
```
--> Missing a comma in the sidebar panel

Question 5
Consider the following code in ui.R
```{r, cache=TRUE}
#   shinyUI(pageWithSidebar(
#   headerPanel("Example plot"),
#   sidebarPanel(
#     sliderInput('mu', 'Guess at the mu',value = 70, min = 60, max = 80, step = 0.05,)
#   ),
#   mainPanel(
#     plotOutput('newHist')
#   )
#   ))
```

And the following server.R code.
```{r, cache=TRUE}
# library(UsingR)
# data(galton)
# 
# shinyServer(
#   function(input, output) {
#     output$myHist <- renderPlot({
#       hist(galton$child, xlab='child height', col='lightblue',main='Histogram')
#       mu <- input$mu
#       lines(c(mu, mu), c(0, 200),col="red",lwd=5)
#       mse <- mean((galton$child - mu)^2)
#       text(63, 150, paste("mu = ", mu))
#       text(63, 140, paste("MSE = ", round(mse, 2)))
#       })
#     
#   }
# )
```

Why isn't it doing what we want?
--> The server.R output name isn't the same as the plotOutput command used in ui.R.

