---
title: "Sample means from an exponential distribution and its behaviour towards CLT"
author: "Jens Schreiber"
date: "24 Mai 2015"
output:
  pdf_document:
    keep_tex: yes
  html_document: default
  word_document: default
---
In this report the exponetial function is investigated and compared to its behaviour towards the Central Limit Theorem (CLT). The CLT states that when we take averages of Independent and Identical Distributed variables, the averages become that of a normal distribution as sample size increases. If we consider $\bar{X}_n$ as the  continuously created sample means of the exponential function then it is approximately $N(\mu,\sigma^2 / n)$ under the CLT, where $\sigma^2 / n$ is the standard error. 

For the simulation to show the behaviour of sample means of the exponetial function
we will calcluate the mean from $n=40$ samples of an exponetial function using the function "rexp" of the R programming langues. This will be done $1000$ times and then its characteristics is investigated. For simulations we use $\lambda=0.2$.

The following R code set-up's the required parameters:
```{r,cache=TRUE}
  lambda = 0.2
  meanTheor  <- 1/lambda
  sdTheor  <- 1/lambda
  varTheor<-sdTheor^2
  n  <- 40
  nrSim <- 1000
```

Please note that for the exponential function the mean is $1/\lambda$ and the standard deviation is $1/\lambda$. 

### Simulation
In order to show the CLT with the exponential function we create $n=40$ samples with $\lambda=0.2$, attach its mean to the sampleMean vector and do this $1000$ times, see the the following code. For convenience we store the data in the sampleData frame in the last step.

```{r}
  set.seed(42)
  sampleMeans = NULL
  for (i in 1 : nrSim) sampleMeans = c(sampleMeans, mean(rexp(n, lambda)))
  sampleData<-data.frame(sampleMeans)
````

### Comparison of theoretical and sample mean

The folloing R code creates an histogram of the simulated data and plot their theoretical (in blue) and their sample mean (in red) with dashed line. 
```{r,cache=TRUE}
  library(ggplot2)
  ggplot(sampleData, aes(x=sampleMeans)) + 
  # Histogram with count on y-axis
  geom_histogram(aes(y = ..density..),  binwidth=.5, colour="black", fill="white") +  
  # line of sample mean
  geom_vline(aes(xintercept=mean(sampleMeans, na.rm=T)),  color="red", 
             linetype="dashed", size=1) +
  # line of theoretical mean
  geom_vline(aes(xintercept=meanTheor),  color="blue", linetype="dashed", size=1)  
```

Their theoretical and sample mean are so close to each other it is hard to distinguish them. The theoretical mean is $5.0$ and depending on the simulation the sample mean is $4.99$, with "set.seed(42)", which is close to the theoretical mean of $5.0$

```{r, cache=TRUE}
  c(meanTheor,mean(sampleData$sampleMeans))
```


### Comparision of theoretical varaince and the variance of the distribution.
The theoretical variance or lets rather say the approximated variance is just the variance of the sample distribution from which we create our sample means divided by the the number of created samples. (Note: It doesn't matter if we do $1000$ simulations it will always be the same value.) So the theoretical variance is given by: 
```{r, cache=TRUE}
  varTheor / n
```

Which is really close to our actual value of our sample means:

```{r, cache=TRUE}
  var(sampleData$sampleMeans)
```

The acutual result depends upon the simulation and will be different for different "set.seed()".


### Test whether the distribution is approximately normal

In this section we want to show that the created distribution of sample means from the exponential fuction is approximately normal. One argument is a comparision of the samples of an exponential function (simulated by "rexp(nrSim,lambda)") and the the sample means of an exponential function. For the latter samples we use the previously simulated data.

```{r,  echo=FALSE}
  set.seed(42)
  par(mfrow = c(1, 2),oma=c(0,0,2,0))
  hist( rexp(nrSim,lambda) ,xlab="Samples of exponential function", main="")                
  hist( sampleData$sampleMeans, xlab="Sample means of exponential function", main="")
```

In the comparison of the two plots, we see that the samples of the exponential function follow obviously an exponential shaped distribution, where the simulated data creates a bell shaped histogramm simular to that of a normal distribution.

Furthermore we can use the following Q-Q-Plot to show its approximal normal distribution:

```{r, cache=TRUE}
  qqnorm(sampleMeans)
  qqline(sampleMeans)
```

This kind of plots, in general, can be used to compare quantiles of different distributions. Here we specifically want to compare the quantiles of the normal distribution to that of our sample mean distribution. With the "qqline" function we add a line to the plot of the theoretical normal distributuon. That's why if the sample mean distribtion is approximated normal distributed all sample quantiles should be situated on the line. 

At least between quantiles -2 and 1 or 1.5 the former mentioned conditions holds true, so we can expect the distribution to be approximately normal distributed. 

<!-- In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials.  You should
1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

In point 3, focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials. 

As a motivating example, compare the distribution of 1000 random uniforms
hist(runif(1000))
and the distribution of 1000 averages of 40 random uniforms

mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(runif(40)))
hist(mns)
This distribution looks far more Gaussian than the original uniform distribution!


This exercise is asking you to use your knowledge of the theory given in class to relate the two distributions.  
Confused?  Try re-watching video lecture 07 for a starter on how to complete this project.


Sample Project Report Structure

Of course, there are multiple ways one could structure a report to address the requirements above.  However, the more clearly you pose and answer each question, the easier it will be for reviewers to clearly identify and evaluate your work. 

A sample set of headings that could be used to guide the creation of your report might be:

Title (give an appropriate title) and Author Name
Overview: In a few (2-3) sentences explain what is going to be reported on.
Simulations: Include English explanations of the simulations you ran, with the accompanying R code. Your explanations should make clear what the R code accomplishes.
Sample Mean versus Theoretical Mean: Include figures with titles. In the figures, highlight the means you are comparing. Include text that explains the figures and what is shown on them, and provides appropriate numbers.
Sample Variance versus Theoretical Variance: Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances.
Distribution: Via figures and text, explain how one can tell the distribution is approximately normal.--> 



