---
title: "Quiz3"
author: "Jens Schreiber"
date: "19. August 2015"
output: html_document
---

Question 1
Load the cell segmentation data from the AppliedPredictiveModeling package using the commands:

```{r, echo=FALSE}
  library(AppliedPredictiveModeling)
  data(segmentationOriginal)
  library(caret)
  library(rpart)
```
1. Subset the data to a training set and testing set based on the Case variable in the data set. 
2. Set the seed to 125 and fit a CART model with the rpart method using all predictor variables and default caret settings. 


```{r, echo=FALSE}
  sessionInfo()
  #head(segmentationOriginal,2)
  names(segmentationOriginal)
  training = segmentationOriginal[segmentationOriginal$Case=="Train", ]
  testing = segmentationOriginal[segmentationOriginal$Case=="Test", ]
  set.seed(125)
  rpartFit<-rpart(Class ~ .,data=training)
  printcp(rpartFit) # display the results 
  plotcp(rpartFit) # visualize cross-validation results 
  #summary(rpartFit) # detailed summary of splits

  # plot tree 
  plot(rpartFit, uniform=TRUE, 
    main="Classification Tree for Case")
  text(rpartFit, use.n=TRUE, all=TRUE, cex=.8)
  newdata.example <- make.obs(c(TotalIntenStatusCh3 = 3, FiberWidthCh1 = 20, VarIntenCh4 = 100))
  predict(rpartFit, newdata=newdata.example)
```

3. In the final model what would be the final model prediction for cases with the following variable values:
a. TotalIntench2 = 23,000; FiberWidthCh1 = 10; PerimStatusCh1=2 
b. TotalIntench2 = 50,000; FiberWidthCh1 = 10;VarIntenCh4 = 100 
c. TotalIntench2 = 57,000; FiberWidthCh1 = 8;VarIntenCh4 = 100 
d. FiberWidthCh1 = 8;VarIntenCh4 = 100; PerimStatusCh1=2 

