---
title: "Influences on MPG in the mtcars dataset"
author: "Jens Schreiber"
date: "17. Mai 2015"
output: html_document
---

This report shall evaluated the influences on "miles per galon (MPG)" in the mtcars dataset. Particullary we are intersted in answering if an automatic or a manual transmission is better for MPG and quantify the MPG difference between a manual and a automatic transmission. Therefore we will do an exploratory analysis and evaluate different regression models.

Initial set-up:
```{r,cache=TRUE}
  library(datasets);library(ggplot2);library(dplyr)
  #assures that factor variables are correctly included as factor variables
  mtcars$am<-as.factor(mtcars$am); mtcars$cyl<-as.factor(mtcars$cyl)
  mtcars$gear<-as.factor(mtcars$gear);mtcars$carb<-as.factor(mtcars$carb)
  mtcars$vs<-as.factor(mtcars$vs)
```

###Exploratory Analysis
At first one might be intersted if the mean MPG for a manual or an automatic transmission is higher. In figure 1 we see that the manual transmission (factor of am=1) has a higher mean value for MPG and also a higher standard deviation. Typically one would expect that also there is a relation between the weight (wt) and mpg as well as between the horse power (hp) and mpg. 
```{r,cache=TRUE}
  cor(mtcars$wt,numeric(mtcars$cyl))
  cor(mtcars$hp,numeric(mtcars$cyl))
```
Due to a correlation between the cylinders (cyl) and wt, as well as between cyl and hp we include cyl as a factor variable in the figures 2 and 3 in the appendix.

###“Is an automatic or manual transmission better for MPG”

In order to answer the question if an automatic or manual transmission is better for MPG one must also include other variables which might influence MPG besides the transmission method. 

```{r,cache=TRUE}
  library(car)
  summary(aov(lm(mpg~.,data=mtcars)))
```

The aov function suggests to include the variables cyl, wt and disp. Based on the exploratory analysis and our particular question regarding the influence of tha am variable we also will include am and hp for our evaluation. 
```{r,cache=TRUE}
  fit1 <- lm(mpg~am,data=mtcars)
  fit2 <- lm(mpg~am+cyl,data=mtcars)
  fit3 <- lm(mpg~am+cyl+disp,data=mtcars)
  fit4 <- lm(mpg~am+cyl+disp+wt,data=mtcars)
  fit5 <- lm(mpg~am+cyl+disp+wt+hp,data=mtcars)
  anova(fit1,fit2,fit3,fit4,fit5)
```

```{r,cache=TRUE}
  c(summary(fit1)$adj.r.squared,summary(fit2)$adj.r.squared,
               summary(fit3)$adj.r.squared,summary(fit4)$adj.r.squared, 
               summary(fit5)$adj.r.squared)
``` 

```{r,cache=TRUE}
    summary(fit1)
    summary(fit5)
``` 

###"Quantify the MPG difference between automatic and manual transmissions"


###Appendix
####Figure 1:
```{r,cache=TRUE,message=FALSE, warning=FALSE,fig.height=5, fig.width = 5}
  ggplot(data=mtcars, aes(x=factor(am), y=mpg,fill=factor(am))) + geom_boxplot()
```

####Figure 2:
```{r,cache=TRUE,fig.height=5, fig.width = 5}
  qplot(y=mpg, x=wt, data=mtcars, shape=factor(cyl), color=factor(cyl),main="", xlab="Weight", ylab="MPG")  
```

####Figure 3:
```{r,cache=TRUE,fig.height=5, fig.width = 5}
  qplot(y=mpg, x=hp, data=mtcars, shape=factor(cyl), color=factor(cyl),main="",xlab="HP", ylab="MPG")  
```

<!--
Did the student interpret the coefficients correctly?
Did the student do some exploratory data analyses?
Did the student fit multiple models and detail their strategy for model selection?
Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?
Did the student do a residual plot and some diagnostics?
Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?


```{r,cache=TRUE}
  ggplot(mtcars, aes(x=wt, y=mpg,color=factor(am))) +
    geom_point(shape=1) +   
    geom_smooth(method=lm)
```

```{r,cache=TRUE}
  ggplot(mtcars, aes(x=hp, y=mpg,color=factor(am))) +
    geom_point(shape=1) +   
    geom_smooth(method=lm)
```

```{r,cache=TRUE}
  ggplot(mtcars, aes(x=cyl, y=mpg,color=factor(am))) +
    geom_point(shape=1) +   
    geom_smooth(method=lm)
```
-->
