---
title: "Influences on MPG in the mtcars dataset"
author: "Jens Schreiber"
date: "24 Mai 2015"
output:
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
  html_document: default
fontsize: 8pt
---
###Executive Summary

* This report shall evaluated the influences on "miles per galon (MPG)" in the mtcars dataset. Particullary we are intersted in answering the following two questions: 
    + Is an automatic or manual transmission better for MPG?
    + What is the MPG difference between automatic and manual transmissions?
* Therefore we will do an exploratory analysis and evaluate different regression models, where MPG is the evaluated outcome.
* During evaluation of different variables for a linear model hp, wt, disp and cyl are the most promissing
* In the final model for predicting the outcome mpg we use the predictors wt,hp and am and obtain an adjusted r squared value of 0.82 
    + (only am as predicor for MPG has an adjusted r squared error of 0.34)
* In the final model we reject that am has significant influence with a p-value of 0.14 and a significance level of 0.05
<!--* However we claim that hp with a p-value of 0.0005464023 and wt with a p-value of  0.003574031 have significant influence on the model--!>
* In conclusion we fail to answer the two questions due to the insignificance of am for predicting MPG in our final model

###Exploratory Analysis

Initial set-up:
```{r,cache=TRUE,message=FALSE, warning=FALSE,comment=FALSE}
  library(datasets);library(ggplot2);library(dplyr)
  #assures that factor variables are correctly included as factor variables
  mtcars1<-mtcars;mtcars1$am<-as.factor(mtcars$am); mtcars1$cyl<-as.factor(mtcars1$cyl)
```

At first one might be intersted if the mean MPG for a manual or an automatic transmission is higher. In figure 1 we see that the manual transmission has a higher mean value for MPG and a higher standard deviation. In the correlation plot of figure 2, we see that mpg has the highest correlation to wt, disp, cyl and hp. Also we see a high correlation between cyl to disp, wt and hp. Furtheremore there is a high correlation between hp to cyl and disp.

###Selection of linear regression models

In order to answer the question if an automatic or manual transmission is better for MPG one must also include other variables which might influence MPG besides the transmission method. 

<!--
```{r,cache=TRUE}
  #mtcars1$gear<-as.factor(mtcars1$gear);mtcars1$carb<-as.factor(mtcars1$carb)
  #mtcars1$vs<-as.factor(mtcars1$vs)
  summary(aov(lm(mpg~.,data=mtcars1)))
```
--!>

Based upon the correlation to mpg and the required analysis of am we analyse the following models:
```{r,cache=TRUE,echo=TRUE}
  fit1 <- lm(mpg~am,data=mtcars1);fit2 <- lm(mpg~am+cyl,data=mtcars1)
  fit3 <- lm(mpg~am+cyl+disp,data=mtcars1);fit4 <- lm(mpg~am+cyl+disp+hp,data=mtcars1)
  fit5 <- lm(mpg~am+cyl+disp+hp+wt,data=mtcars1);fit6 <- lm(mpg~am+hp+wt,data=mtcars1)
  anova(fit1,fit2,fit3,fit4,fit5,fit6)
```
For the models fit2,fit3,fit4 and fit4 the p-value is below 0.05 so we reject $H_0$ in favour of $H_1$ so that there is significant difference to our base model of fit1. However for fit6 we fail to reject $H_0$, so that there is no significant difference between fit5 and fit6. This can be confirmed by the adjusted r squared error. The difference in the adjusted r squared error between fit5 and fit6 is around 0.012, so there is no major improve. This is due to the correlation between hp to disp and cyl, which was shown in the exploratory analysis section. 
```{r,cache=TRUE}
  c(summary(fit1)$adj.r.squared,summary(fit2)$adj.r.squared,summary(fit3)$adj.r.squared,
    summary(fit4)$adj.r.squared, summary(fit5)$adj.r.squared,summary(fit6)$adj.r.squared)
``` 

The model of fit6 agrees with our natural understanding that horse power and the weight of a car mainly influence MPG. In the Q-Q-Plot of figure 3 we see that the residuals seem to be normally distributed. In the "Residuals vs Fitted"-Plot there is no clear pattern, even though we see more negative values which menas our prediction was too high. The two former arguments strengths the believe in our previous results. 
 

###Quantification of the MPG difference between automatic and manual transmissions
In the summary of the most reasonable fit (fit6), the p-value for am is aroud 0.141, in case of a significance level of 0.05 we would reject $H_1$ in favour of $H_0$. So, there is no significance influence of the transmission type. Based on the small p-values for hp and wt, those seem to be the biggest influence for MPG with an mpg offset of 34  a negative slope of -0.038 for hp and a slope of -2.88 for wt. Due to the insignificance of am we cannot quantify the difference between automatic and manual transmissions for the model of fit6.
```{r,cache=TRUE,echo=TRUE}
    summary(fit6)$coefficients
```



##Appendix

####Figure 1:

```{r,cache=TRUE,message=FALSE, warning=FALSE,fig.height=5, fig.width = 7}
  ggplot(data=mtcars1, aes(x=am, y=mpg,fill=am),xlab="Transmission Type",ylab="MPG") + 
  geom_boxplot() + scale_fill_discrete(name="Transmission Type (am)",
                                       labels=c("Automatic", "Manual"))
```

\newpage

####Figure 2:


```{r,cache=TRUE,fig.height=5, fig.width = 5}
  library(corrplot)
  corrplot.mixed(cor(mtcars), order = "AOE")
```
\newpage

####Figure 3:


```{r,cache=TRUE,fig.height=5, fig.width = 5}
  par(mfrow=c(2,2))
  plot(fit6)
```

<!--
Did the student interpret the coefficients correctly?
Did the student do some exploratory data analyses?
Did the student fit multiple models and detail their strategy for model selection?
Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?
Did the student do a residual plot and some diagnostics?
Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?


```{r,cache=TRUE}
  ggplot(mtcars, aes(x=wt, y=mpg,color=factor(am))) +
    geom_point(shape=1) +   
    geom_smooth(method=lm)
```

```{r,cache=TRUE}
  ggplot(mtcars, aes(x=hp, y=mpg,color=factor(am))) +
    geom_point(shape=1) +   
    geom_smooth(method=lm)
```

```{r,cache=TRUE}
  ggplot(mtcars, aes(x=cyl, y=mpg,color=factor(am))) +
    geom_point(shape=1) +   
    geom_smooth(method=lm)
```
-->
